# INTRODUCTION {#intrap}

## Aim

* Principal: 
    - Review the methods previously applied to the standarization across ELISA plates 
    in vaccine or serological surveillance research.

* Secondary: 
    - Register project updates, current problems to solve, project organization 
    and benefits of using Git.

## To Do

## Dependencies

No extra package required.

```{r, results='hide', message=FALSE}
##essential
library(XLConnect)    # load EXCEL workbooks
library(drc)          # Dose-Response modeling
##accesory
library("DiagrammeR") # method Flowchart
library("knitr")      # To display nice tables
library("tidyverse")
#ggplot2, tibble, tidyr, readr, purr, dplyyr
##extra
library("Rmisc")      # multiplot ggplots
library("haven")      # import STATA df
library("readxl")     # import EXCEL df
library("forcats")    # factor vectors
library("viridis")    # color visualization
```

## Input
```{r}
Pf.ID <- readRDS("data/01-dilution.rds")
#unlist(Pf.ID[1:12,])
```


## APPENDIX: From OD to Ab units

- Sepulveda et.al. [@Seplveda2015] cite two main publications:
    1. **Cook et.al., 2011 in Bioko Island, Equatorial Guinea** [@Cook2011], and
    2. **Cunha et.al., 2014 in Jacareacanga, Brazil** [@Cunha2014].
- Both papers followed the same experimental procedure to standardize the data 
across plates. 
Althought, non of them cite **Miura et.al., 2008**[@Miura2008].
- Nuno Sepulveda is co-author of the second paper [@Cunha2014].


### NAMRU-6 method

```{r}
t<-seq(1,12,1)
x <- c()
x0 <- 50
f <- function(x){x*2}
x[1]<-x0
for (i in 1:(length(t)-1)) {
  x[i+1] = f(x[i])
}
```


The method provided by Julio followed this procedure, making two main assumptions:

1. An **endpoint titer** equal to the last dilution of the STD, in this case `r format(x[length(x)], scientific=FALSE)`.
2. An equal **endpoint titer** among *P.falciparum* and *P.vivax* STD's.

```{r}
example <- cbind(unlist(Pf.ID[1:12,]), 
                 x, 
                 gsub('STD 1','102400',unlist(Pf.ID[1:12,])),
                 x[length(x)]/x)
colnames(example) <- c("Dilution", "Reciprocal of Dilution", "Factor", "ELISA Ab units")
kable(example)
```


### K.Miura method

K.Miura et.al. [@Miura2008] previously assigned Antibody Units to the STD 
(as the reciprocal dilution giving an OD~450nm~=1*). After this step, the STD 
was applied to each ELISA plate.

  (@) For example, for the **Anti-Pvs25 mokey sera**, after generating aliquotes 
  of 1:100 from the initial pool, four-fold serial dilution was perfomed. Here, 
  an OD~450nm~=1 had a Rec.Dilution of 20,000. Then STD Ab units was equal to 20,000. 

  (@) After the assignation of Ab units to the STD, a two-fold serial dilution starting 
  with a dilution of 20 Ab units, equivalent to 1:1000 dilution, was applied on each 
  ELISA plate.

  (@) As an example, in the paper is said that the 5th dilution of the STD sera 
  is suppoused to have 1.25 Ab units. 
  So, if $1.25=\frac{STD}{16000}$ then the STD Ab units are equal to 20,000.

*The reason behind an OD~450nm~=1 may be in the variability of the STD curve Upper 
and Lower limits. This behaviour also is observed in the comparoson of the STD OD 
values across ELISA plates.

### Corran method

Cook et.al.[@Cook2011] cite Corran et.al. [@Corran2008] which have a full description 
of the methods.

Corran et.al.[@Corran2008] make the assumption of an undiluted (initial) concentration 
of the sera pool previous to a 3pLL model fitting. By this way, the OD is directly 
proportional to the independent variable. 
[suppl.file 1](https://static-content.springer.com/esm/art%3A10.1186%2F1475-2875-7-195/MediaObjects/12936_2008_670_MOESM1_ESM.pdf)

At Methods, they write the following:

> "A titration curve was fitted to the ODs obtained for the standard plasma dilutions 
> by least squares minimisation using a three variable sigmoid model and the solver 
> add-in in Excel (Microsoft), assuming an arbitrary value of **1000 Units/ml** of 
> antibody against each antigen in the **standard pool**. OD values for the spot 
> extracts were converted to units/ml using this fitted curve."

***

## APPENDIX: the 4pLL model

The **four-parameter log-logistic function** have the following presentations [@weimer2012impact]: 
$$f(x)=f(x;b,c,d,e)=c+\frac{d-c}{1+\exp[b(log(x)-log(e))]}=c+\frac{d-c}{1+\left(\frac{x}{e}\right)^b}=c+\frac{d-c}{1+{10}^{b(log_{10}(x)-log_{10}(e))}}$$

By isolating $x$, equivalent expressions are commonly reported as inverse functions 
of the 4pLL equation, assuming `b=-1` and then `c=0`:
$$x=e\left[{\left(\frac{d-f(x)}{f(x)-c}\right)}^{1/b}\right] \qquad \Rightarrow \qquad x=\frac{e}{\frac{d-c}{f(x)-c}-1} \qquad  \Rightarrow \qquad x=\frac{e}{\frac{d}{f(x)}-1}$$

If no log-transformation is required, then the following expression are often employed, 
depending on the scale of `x`: 
$$f(x)=c+\frac{d-c}{1+\exp[b(x-e)]} \qquad or \qquad f(x)=c+\frac{d-c}{1+{10}^{b(x-e)}}$$ 
For example, this last expression, with `x` in decimal-log scale, is present in the 
page 7 of the *drLumi* [@drLumi] package vignette.

Lastly, when using the nls function to fit the data, 
[Self-Starter](https://socserv.socsci.mcmaster.ca/jfox/Books/Companion/appendix/Appendix-Nonlinear-Regression.pdf) functions are required. 
This functions, in contrast with the last example, requires `x` in natural-log scale. 
E.g. the logistic model have the following parameters 
$$f(x)=\frac{\phi_1}{1+\exp\left[ \frac{\phi_2-x}{\phi_3}\right]} \qquad \Rightarrow \qquad f(x)=0+\frac{\phi_1-0}{1+\exp\left[ \frac{-1}{\phi_3} (x-\phi_2) \right]} $$

### Publications

  1. **Miura et.al, 2008** [@Miura2008] reported a "four-parameter hyperbolic curve" in this form: 
  $$x=e\left[{\left(\frac{d-f(x)}{f(x)-c}\right)}^{1/b}\right]$$ 
  where $x={Ab}_{units}$ and $f(x)={OD}_{450nm}$ without parameter letters definition. 
  Those parameter letters have been changed in order to keep concordance with the 
  extended description gave [above](#method).

  1. **Cook et.al., 2011 in Bioko Island, Equatorial Guinea** [@Cook2011] cite 
  Corran et.al. [@Corran2008] which have a full description of the methods 
  [suppl.file 1](https://static-content.springer.com/esm/art%3A10.1186%2F1475-2875-7-195/MediaObjects/12936_2008_670_MOESM1_ESM.pdf). 
  Briefly, they report a ligand-binding equation, with the following inverse function: 
  $$A=\frac{A_{max}*B}{B+k} \qquad \Rightarrow \qquad B=\frac{C}{D}=\frac{k}{\frac{A_{max}}{A}-1}$$ 
  That equation takes a recognizable form after a rearrangement, showing a 4pLL assuming `c=0` and `b=-1`, the same as with the inverse form: 
  $$A=0+\frac{A_{max}-0}{1+\left(\frac{B}{k}\right)^{-1}} \equiv f(x)=0+\frac{d-0}{1+\left(\frac{x}{e}\right)^b}  \qquad and \qquad B=\frac{k}{\frac{A_{max}-0}{A-0}-1} \equiv x=\frac{e}{\frac{d-0}{f(x)-0}-1} $$
    
  2. **Cunha & Sepulveda et.al., 2014 in Jacareacanga, Brazil** [@Cunha2014] report 
  a formula without any details, as in the cited publication [@Bousema2010]: 
  $$ {titer}= \frac{dilution}{ \frac{maximum.OD}{OD.test.serum - minimum.OD} -1} \qquad \neq \qquad x=\frac{e}{\frac{d-c}{f(x)-c}-1} $$ 
  Eventually, they tried to write the inverse of the 4pLL model but fail in the 
  description of the variables.
  

**CONCLUSION:**

Here I have shortly compare the equivalent expressions applied by several reference 
publications [@Miura2008; @Cook2011; @Cunha2014; @Corran2008], all equal to the 
**four-parameter log-logistic function** applied in this report.

***


## APPENDIX: Updates

- **From 09dic2016 to 01feb2017**

*09dic2016:*

With the Antibody Units:

- Conversion of the Estimated Reciprocal Dilutions to *Arbitrary Ab units*.
    + *SOL:* Followed NAMRU-6 SOP for ELISA data analysis.
    + A complementary description is in the [Appendix](#appendix-from-od-to-ab-units)

- Complete understanding of Cook et.al. [@Cook2011] and Cunha et.al. [@Cunha2014] or 
Cook et.al. [@Cook2011] methodology.
    + *SOL:* A complementary description is in the [Appendix](#appendix-the-4pll-model).

- Blank wells were added as STD with Ab units equal to zero. Then, a 4pLL was the 
best model to fit STD data.
    + *NOTE:* This allows to avoid background subtraction.
    
- Positive Control added to graph and Dose estimation
    + *NOTE:* ctrl+ out of STD range should be discussed.
    
*16dic2016:*

- A comparison between both states of Background Subtraction [ON] vs [OFF].
    + *SOL:* An [OFF] state after a Box-Cox transformation results in lower amount 
    of NaN's among the first 5 plates.

- Addition of residual variance and QQ-plots before and after Box-Cox transformation.

- Histogram, Density plots and QQ-plots for OD vs Ab.unit comparison.
    + *NOTE:* No dramatic changes after Background Subtraction.

*23dic2016:*

- 2 main outputs:
    + 01 SIMPLE I/O report for all ELISA templates (w/o Background subtraction)
    + 01 EXTENDED report per ELISA template available if required

- Flowchart of methodology

- 04 Templates with lower samples than expected:
    + N12 (n=39), N18 (n=37), N23 (n=10), N41 (n=18).

- 01 Template with more samples than expected:
    + N15 (n=41)
        - **ZG181-1** sampled against **Pviv** plate only
        - **ZG182-1** sampled against **Pfal** plate only.
    + Both generated NaNs at the end. Needs to be solved.

- Some XLS file modifications:
    + from N20 to N37 and from N39 to N41, 03 extra rows after "Plate 1 QTSR Pf" 
    were deleted.
    + N39 typo at cell M36 corrected.
    + N18 typo at cell J30 corrected.
    + from N23 to N26, STD cell format changed from 'character' to 'number'.
    + N41 blank and ctr+/- were moved to the right position.
    
*30dic2016:*

- Unbiased selection of Ab.units estimate among replicated samples:
    + *NOTE:* This assumes that no dilution to the original evaluated sample has 
    been performed.
  
- About **73 NaN** Ab.units, since only one is above its model Upper Limit, I suggest 
to avoid all this measurement.
    + Requires discussion!
    
*06ene2017:*

- Improvement of the unbiased selection criteria:
    + For triplicates, the closest value to the mean among replicates was selected
    + Reads with mean.OD with a %CV higher than 20% were rejected
    + For duplicates, the replicate with lower %CV was preferred.
    + If a selection was not acomplished, the last measurement was preferred.
    + Criteria was implemented in a function called `Unbiased()`

- Ab.units data frame filtering:
    + Retrieve of all replicated reads: triplicates, duplicates and NaN.
    + Addition of selected reads among triplicates and duplicates.
    + Visualization of mean.OD-Ab.unit non-linearity and whole %CV before and 
    after the procedure

- Brefly: 
    + 964 samples are **in** with 02 reads each, 01 per specie
    + 49 samples are **in** with only 01 read each
    + 03 samples are **out** because of NaN on both reads

- From the resting 55 reads with unestimated Ab.unit:
    + Only 03 reads have a mean.OD **above** their respective model upper limit
        + a re-run may be discussed.
    + All the rest, have a mean.OD **below** the lower limit
        + These reads aproaches to zero.

- Done with 02 main outputs:
    + a SIMPLE **.html** report to share with collaborators.
    + a FINAL **.csv** with the obtained Ab.units with the 
    *Zungarococha_database1_3Oct2016.csv* file format
        + Better suited for **covariates** addition.

- For a comparison between resting NaN's mean.OD and Upper/Lower limits per ELISA plate:
    + Complete list of NaN's available in 
    [02-dtfilter.html](#filter)
    + Complete list of Upper/Lower limits per ELISA plate in 
    [01-standard.html](#stdrad)

- A complete list of previous updates available in [Appendix](#appendix-updates)

*13ene2017:*

- **To discuss:** Criteria for NaN:
    + If approaching zero, then change NaN to 0? (n=46+6)
    + If aproaching infinity, then dilute and re-run? (n=3)
        + Review Corran, et al. criteria.
    
- The original covariate DB is required to fill then into the final DB
    + `Zungarococha_database1_30oct2016` have only **960 samples**
    + There is no data for **09 samples**, and
    + There are **02 samples** that are not present in the 28 evaluated Templates.

*20ene2017:*

- After Katty indication, DB typos were corrected in the current script at 
[02-dtfilter.html](#filter):
    + 2 samples were added: ZG009-3 and NN068-2.

- Complete list of replicates before and after the unbiased selection available at 
[02-dtfilter.html](#filter)

- Covariate DB was added to the analysis at 
[03-covariat.html](#covar):
    + Updated covariate DB was merged with the Ab.units per sample:
        + CSV file is available to re-analyse seropositivity in STATA.
    
    + Contrast of covariates and Ab.units between patients with doble ID?
        + Based on the current data, there is not enough support to delete those samples.

    + A comparison between covariate DB PRE and POST update is also available
        + Frequencies and age distribution per community. 

- **To discuss:** Criteria for NaN:
    + If approaching zero, then change NaN to 0? (n=46+6)
    + If aproaching infinity, then dilute and re-run? (n=3)

*01feb2017:*

- **To discuss:** Criteria for NaN:
    + If approaching zero, then change NaN to 0? (n=46+6)
    + If aproaching infinity, then dilute and re-run? (n=3)

- Covariate [labels](#new-db-stats) could be changes if required.

- Covariates visualization of the 03 samples requiring a [re-run](#re-run).

- Project folder arrangement ready to share via *Git(Hub)*[@CienciaReproducible2016].

**NEXT STEPS:**

- Writing of methodology and its limitations for a potential publication.

- Replicate seropositive analysis.
    + Check if "NaN" data cells creates dificulties. (easy to correct)
    + Contrast mean.OD and standardized Ab.units serpositivity patters per community.

- Share code via [Git(Hub)](https://speakerdeck.com/alicebartlett/git-for-humans) 
with Nuno.
    + BENEFITS:
        + Will improve collaboration:
            + code review, version control, error tracking and paper writing.
        + Will improve publication process: 
            + journal peer-review process, preference on 
            [plain-text manustripts](https://github.com/rstudio/rticles) 
            with code available.
        + Will improve publication quality and impact:
            + sharing a reproducible workflow acelerates scientific progress
                + Remember that standardized open-access procedures are still a 
                deficit in this field.
                + Recent "open-acces" publications by Biggs [@biggs2017] and 
                Dewasurendra [@dewasurendra2017effectiveness] *do not* provide 
                open-acces methods 
                + only in the last [@dewasurendra2017effectiveness] specifies that 
                "datasets and scripts are available under reasonable request".
            + This type of reports, relevant for malaria surveillance 
            [@elliott2014surveillance], would be highly improved with standardize 
            open-acess lab/comp protocols.
        + Will ease the transition from the Rproject to a Rpackage aim to analyse 
        this specific data.
            + Nuno have already made most of the require methods [@Seplveda2015] 
            and simulations [@sepulveda2015sample] with open-access tools.

**PROBLEMS TO BE SOLVED:**

- *MUST:* Improve code
    + Replace SDCV by an extended MEAN data.frame.
    + This will solve N15 NaN problem



## APPENDIX: Version Control

- **From 30jan2017 to 06mar2017**

- Some link of interest
    + [Git for humans](https://speakerdeck.com/alicebartlett/git-for-humans)
    + [Git and Github with Rstudio](http://r-pkgs.had.co.nz/git.html)

### Project directory

- Following Reproducible Research workflow 
[procedures](http://blog.jom.link/implementation_basic_reproductible_workflow.html)

```{bash}
less notrack/project_elixr.md
```

### Git version control

- For a graphical introduction to Git within Rstudio, check this 
[link](http://r-pkgs.had.co.nz/git.html).

```{bash}
git log --oneline
```

- For a restricted file tracking, e.g. avoiding large files like datasets, 
add their names to a `.gitignore` file
```{bash}
less .gitignore
```


### Git commits in detail

```{bash}
git log
```


## APPENDIX: README file

```{bash}
less README.md
```


<!--## References-->