---
title: "CAPI vs PAPI"
author: "Andree Valle Campos"
date: '`r Sys.Date()`'
output:
  html_notebook:
  #html_document:
    fig_caption: yes
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
    code_folding: "hide"
    #df_print: paged
#bibliography: malaria.bib
biblio-style: apalike
link-citations: yes
#csl: american-medical-association.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, #fig.path = "01-",
                      warning = FALSE)
#knitr::opts_knit$set(root.dir = '../.')

options(width = 90) # expand limits of CONSOLE output

#require(Hmisc)
#mu <- markupSpecs$html   # markupSpecs is in Hmisc
# hidden command (<code>r mu$widescreen()</code>) to use an entire wide screen. # mu$widescreen()
#`r mu$widescreen()`
```

# CAPI vs PAPI {#capipapi}

## Aim

- Error rate between data input methods.

## Results

## To Do

## Dependencies

```{r, results='hide', message=FALSE}
library(tidyverse)
```

## Input

### Dictionary

```{r}
dic <- readxl::read_xlsx("data-raw/UNAP_CAPIvPAPI_barridoI_2017.xlsx",sheet = 3,col_names = FALSE) %>% 
  select("var"=X__1,"code"=X__3) %>% 
  mutate_all(funs(stringr::str_to_lower(.)))
dic
```

### Data

```{r}
capi <- readxl::read_xlsx("data-raw/UNAP_CAPIvPAPI_barridoI_2017.xlsx",sheet = 1) %>% 
  mutate(B001= stringr::str_replace_all(B001, "_","-")) %>% 
  rename_all(funs(stringr::str_to_lower(.))) %>% 
  mutate_all(funs(stringr::str_to_lower(.))) %>% 
  mutate(b003=stringr::str_replace(b003,"ni.+","nina rumi")) %>% 
  mutate(b002=stringr::str_replace(b002,"(\\d+)/(\\d+)/(\\d+)(.+)","\\3-\\2-\\1"),
         b010=stringr::str_replace(b010,"(\\d+)/(\\d+)/(\\d+)","\\3-\\2-\\1")) %>% #dplyr::count(b100,b095)
  mutate_at(vars(b012:b102),funs(stringr::str_replace(.,"(\\d+)(.+)","\\1"))) #%>% dplyr::count(b100,b095)
papi <- readxl::read_xlsx("data-raw/UNAP_CAPIvPAPI_barridoI_2017.xlsx",sheet = 2) %>% 
  rename_all(funs(stringr::str_to_lower(.))) %>% 
  mutate_all(funs(stringr::str_to_lower(.))) %>% #dplyr::count(b073)
  mutate_at(vars(b012:b102),funs(stringr::str_replace(.,"(\\d+)(.+)","\\1"))) %>% #dplyr::count(b003) #dplyr::count(b073)
  mutate(b003=stringr::str_replace(b003,"ni.+","nina rumi"),
         b003=stringr::str_replace(b003,"lla.+","llanchama"),
         b003=stringr::str_replace(b003,"zu.+","zungarococha")) #%>% dplyr::count(b003)

mix <- capi %>%
  full_join(papi,
            by="b001", suffix=c(".c",".p"))

#mix %>% glimpse()
```

```{r, eval=FALSE}
# ¿HOW TO DO IT PROGRAMMATICALLY?
mix %>%
  select(b003.c,b003.p) %>% 
  mutate(test= b003.c==b003.p) %>% 
  count(test) %>%
  spread(test,n)
```

## Results

### Error in percentage

- __A.all:__ total number of values
- __N.cap__: non-missing values in CAPI
- __N.pap:__ non-missing values in PAPI
- __T.cvp:__ number of TRUE values to test of equality between CAPI vs PAPI data
- __F.cvp:__ number of FALSE to test of equality between CAPI vs PAPI data
- __na:__ non-available data in either or both datasets to perform a test comparison
- __(T/N)%:__ percentage of TRUE among non-missing values
- __(F/N)%:__ percentage of FALSE among non-missing values
- __(na/A)%:__ percentage of NA among all values
- __high.N:__ higher number of non-missing values in either CAPI (1) or PAPI (0)

```{r}
var <- NULL

for(i in 2:dim(capi)[2]){
  test <- capi[,i] == papi[,i] #TEST equality
  a <- length(test)
  a.c <- length(which(capi[,i] != "NA")) #NON-MISSING VALUES in CAPI
  a.p <- length(which(papi[,i] != "NA")) #NON-MISSING VALUES in PAPI
  b <- length(which(test == TRUE))
  c <- length(which(test == FALSE))
  d <- a - (b + c)
  e <- as.numeric(format((b/(b + c))*100,digits = 3)) #proportion among NON-MISSING VALUES
  f <- as.numeric(format((c/(b + c))*100,digits = 1)) #proportion among NON-MISSING VALUES
  g <- as.numeric(format((d/(a))*100,digits = 1)) #proportion of NA among ALL
  h <- ifelse(a.c>a.p,1,0)
  var <- rbind(var,cbind(a,a.c,a.p,b,c,d,e,f,g,h))
  rownames(var)[i-1] <- colnames(capi[,i])
}

colnames(var) <- c("A.all","N.cap","N.pap","T.cvp","F.cvp","na","(T/N)%","(F/N)%","(na/A)%","high.N")

dim(var)
var
```

## Tidy requirement

### 100% NaN values

- retired due to non-equality between variable data as a consequence of _NA_ missing values.
- involved variables: "email" and "otros_"

```{r}
var[var[, "(F/N)%"] == "NaN",] #5

x <- rownames(var[var[, "(F/N)%"] == "NaN",])
length(x)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.))))
#https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r
```


### 100% false among non-missing

- lower amount of non-missing values
- involved variables: "telefono", "donde_", "otro_", "especifique_"

```{r}
var[var[, "(F/N)%"] == "100",] #4

x <- rownames(var[var[, "(F/N)%"] == "100",])
length(x)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.))))
#https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r
```

- b007: cadenas separadas
- b041: orden de cadenas aditivas
- b068: mismo significado, diferente significante
- b078: mismo significado, diferente significante

### 0% false among non-missing

- AQUI: rescatar variables con data relevante

```{r}
var[var[, "(F/N)%"] == "0",] #13
var[var[, "(F/N)%"] == "0" & var[, "na"] != "395" & var[, "(na/A)%"] != "99"  & var[, "(na/A)%"] != "96",] #8
var[var[, "(F/N)%"] == "0" & var[, "(na/A)%"] != "0" & var[, "(na/A)%"] != "100",] #8
var[var[, "(F/N)%"] == "0" & var[, "(na/A)%"] != "0" & var[, "(na/A)%"] == "100",] #4

x <- rownames(var[var[, "(F/N)%"] == "0",])
length(x)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.))))
#https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r
```

### at least one false

- AIM:
    + Check the degree of error
    + If it is due to format, then tidy-it-up and re-run

```{r}
var[var[, "(F/N)%"] != "100" & var[, "(F/N)%"] != "NaN" & var[, "(F/N)%"] != "0",] #79

x <- rownames(var[var[, "(F/N)%"] != "100" & var[, "(F/N)%"] != "NaN" & var[, "(F/N)%"] != "0",])
length(x)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.))))
#https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r
```

### explicit categories

```{r}
dic %>% filter(code=="b003"|code=="b004"|code=="b005"|code=="b006")

expl <- capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.)))) %>% 
  select(2:15) 
expl #mutate_all(funs(as.factor(.))) %>% summary()

expl %>% dplyr::count(b003.c,b003.p) %>% mutate(test= b003.c==b003.p)
expl %>% dplyr::count(b004.c,b004.p) %>% mutate(test= b004.c==b004.p)
expl %>% dplyr::count(b005.c,b005.p) %>% mutate(test= b005.c==b005.p)
expl %>% dplyr::count(b006.c,b006.p) %>% mutate(test= b006.c==b006.p)
```

```{r}
cat <- expl %>% select(-starts_with("b003"),-starts_with("b004"),-starts_with("b005"),-starts_with("b006"))

z <- cat %>% 
  select(ends_with(".c")) %>% 
  rename_all(funs(stringr::str_replace_all(.,".c",""))) %>% colnames()

var[z,]

var[z,7:10] %>% summary()
#w <- var[y,] %>% rownames()
#var[y,] %>% as.tibble() %>% mutate(var= w) %>% select(var,everything())
```

### coded categories

```{r}
num <- capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.)))) %>% 
  select(16:159,-starts_with("b016"),-starts_with("b044"),-starts_with("b057"))

num
```

```{r}
y <- num %>% 
  select(ends_with(".c")) %>% 
  rename_all(funs(stringr::str_replace_all(.,".c",""))) %>% colnames()

var[y,]

var[y,7:10] %>% summary()
#w <- var[y,] %>% rownames()
#var[y,] %>% as.tibble() %>% mutate(var= w) %>% select(var,everything())
```


#### special

```{r}
dic %>% filter(code=="b016"|code=="b044"|code=="b057")

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.)))) %>% 
  select(starts_with("b016"),starts_with("b044"),starts_with("b057"))
```

### final

- add 0% false (done in `k`) + low  NA

```{r}
k <- rownames(var[var[, "(F/N)%"] == "0" & var[, "na"] != "395" & var[, "(na/A)%"] != "99"  & var[, "(na/A)%"] != "96",])
```

- bind all

```{r}
x <- c(z,y,k)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

fin <- var[c(z,y,k),]
fin
fin %>% as.tibble() %>% mutate(high.N=as.factor(high.N)) %>% select(7:10) %>% summary()
```

#### error higher than 20%

```{r}
fin[fin[, "(F/N)%"] >= 10,] #5
#fin[fin[, "(F/N)%"] >= 20,] #5

x <- rownames(fin[fin[, "(F/N)%"] >= 20,])
length(x)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.))))
#https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r
```

#### na higher than 20%

```{r}
fin[fin[, "(na/A)%"] >= 10,] #5
#fin[fin[, "(F/N)%"] >= 20,] #5

x <- rownames(fin[fin[, "(na/A)%"] >= 20,])
length(x)

m <- NULL
for(i in 1:length(x)){
  n <- dic %>% filter(code==x[i]) %>% as.matrix()
  m <- rbind(m,n)
}
m

capi %>% select(1,x) %>% 
  full_join(
    papi %>% select(1,x),
    by="b001",suffix=c(".c",".p")
    ) %>% 
  select(noquote(order(colnames(.))))
#https://stackoverflow.com/questions/29873293/dply-order-columns-alphabetically-in-r
```
