# ELISA STANDARDIZATION {#stdrad}

## Aim

* Principal: 
    - Implement a reproducible standardization of OD values across ELISA plates following 
    **Miura et.al., 2008**[@Miura2008].

* Secondary: 
    - Generate useful outputs to compare standardization quality.


## To Do

- Implement `plater` and `readxl` data input
- Assess the **inter-plate-CV** comparing STD variability across plates to generate a 
**Limit of Quantification**
    + Actual CV is the **intra-plate-CV**
    + Generate a tibble of all STD dilutions OD's
    + Extend `std.par` output to all parameters
- Other inter-plate measurements:
    + blank, positive and negative controls


## Dependencies

The required R packages for this analysis are:

+ `plater` [@plater]
+ `readxl` [@readxl]
+ `tidyxl` [@tidyxl]
+ `drc` [@Ritz2015]

```{r, results='hide', message=FALSE}
##essential
library(tidyverse)    # set of tidy packages (readr, tibble, dplyr, tidyr, ggplot2, ¿purr?)
library(tidyxl)       # read untidy excel formats
library(readxl)       # read excel files as tidy tables
library(plater)       # read, tidy, display microtiter plates
library(drc)          # fit dose-response models
##accesory
library(DiagrammeR)   # create flowchart
```

```{r}
theme_set(theme_bw())
```


## Method

- using `DiagrammeR` [@DiagrammeR]

```{r, fig.align='center', fig.width=9}
#install.packages("DiagrammeR")
#library(DiagrammeR)

DiagrammeR("
  graph LR
    A[XLS<br>data] -.-> |<center>readxl<br>+tidyxl</center>| B{<center>CSV<br>data<br>w/<br>PLATER<br>format</center>}
    B -.-> |plater| C{<center>TIDY<br>data</center>}
    C --> C1[STD]
    C --> E[ctr +/-]
    C --> D1[UNK]
    
    C1 -.-> |drc| C2[4pLL model]
    C2 --> C3[Box-Cox]
    C3 --> F{<center>UNK<br>Ab.units</center>}

    D1 --> D2[mean.OD]
    D2 --> D3[OD %CV]
    D3 --> F
    
    F --> G1[Histogram]
    F --> G2[Density]
    F --> G3[QQPlot]

    style C fill:#ffffff, stroke:#000000, stroke-width:2px    
    style A fill:#ffffff, stroke:#000000, stroke-width:2px
    style B fill:#ffffff, stroke:#000000, stroke-width:2px
    style C1 fill:#ffffff, stroke:#000000, stroke-width:2px
    style C2 fill:#ffffff, stroke:#000000, stroke-width:2px
    style C3 fill:#ffffff, stroke:#000000, stroke-width:2px
    style D1 fill:#ffffff, stroke:#000000, stroke-width:2px
    style D2 fill:#ffffff, stroke:#000000, stroke-width:2px
    style D3 fill:#ffffff, stroke:#000000, stroke-width:2px
    style E fill:#ffffff, stroke:#000000, stroke-width:2px
    style F fill:#ffffff, stroke:#000000, stroke-width:2px
    style G1 fill:#ffffff, stroke:#000000, stroke-width:2px
    style G2 fill:#ffffff, stroke:#000000, stroke-width:2px
    style G3 fill:#ffffff, stroke:#000000, stroke-width:2px
")
```


### Log-Logistic (4pLL) model

The curve of the **log-logistic symetric** model describe the *response* `f(x)` dependent of the *dose* `x` 
and **04 parameters**: $$ f(x)=f(x;b,c,d,e)=c+\frac{d-c}{1+\exp[b(log(x)-log(e))]}\ $$ where: 
  
- `c` is the **lower limit** of the response when the *dose* `x` approaches infinity, 
- `d` is the **upper limit** when the *dose* `x` approaches zero,
- `b` is the **slope** around the **point of inflection**, represented by 
- `e` defined as **effective dose** and commmonly denoted as [@Ritz2015]:
    + `ED50`, `EC50` or `IC50` for continuous responses,
    + `LD50` or `LC50` for binomial responses, and
    + $T_{50}$ for event-time responses.

## plater

```{r,eval=FALSE,echo=FALSE}
a <- "data/example-1.csv"
b <- read.csv(a,header = F) 
#b %>% View
plater::check_plater_format(a)
c <- read_plate(a)
str(c)
```

```{r,eval=FALSE,echo=FALSE}
a <- "data/example-2-part-A.csv"
d <- "data/example-2-part-B.csv"
b <- read.csv(a,header = T) 
e <- read.csv(d,header = F)
#b %>% View; e %>% View
#plater::check_plater_format(a)
plater::check_plater_format(d)
c <- add_plate(data=b,file=d,well_ids_column = "Wells")
str(c)
view_plate(data = c, well_ids_column = "Wells", 
  columns_to_display = c("Concentration", "Killing"))
```

```{r,eval=FALSE,echo=FALSE}
a <- "data/example-1.csv"
b <- "data/more-bacteria.csv"
c <- read.csv(a,header = F) 
d <- read.csv(b,header = F) 
#c %>% View; d %>% View
plater::check_plater_format(a)
plater::check_plater_format(b)
c <- read_plates(files=c(a,b))
str(c)
c %>% dplyr::count(Plate)
```

## Procedure

### generate list

```{r}
# lista de muestras en el directorio
smp <- list.files("data-raw/") #length(smp)
#smp %>% as.tibble() %>% rename("sample"=value)

# número de muestras por comunidad
tmp <- smp %>% as.tibble() %>% dplyr::rename("sample"=value) %>% 
  mutate(template=stringr::str_replace(sample,"(Template_ELISA).+","ok")) %>% 
  filter(template=="ok")
#tmp$sample

# loop para generar rutas
allxls <-  NULL
for(i in 1:length(tmp$sample)){
  #a <- list.files(paste0("data-raw/",tmp$sample[i]))
  b <- paste0("data-raw/",tmp$sample[i]#,"/",a
              )
  allxls <- c(allxls,b)
}
```

### write_plate

```{r}
write_plate <- function(path,sheet,id_range,od_range,more_id,id) {
  f1 <- readxl::read_xlsx(path,range = id_range,
                        sheet = sheet,col_names = F) %>% as.matrix()
  f2 <- readxl::read_xlsx(path,range = od_range,
                        sheet = sheet,col_names = F) %>% as.matrix()
  
  f1[1,1] <- "id"; f2[1,1] <- "od"; f2[1,-1] <- 1:12
  
  f <- rbind(f1[1,],cbind(f1[-1,1],matrix(id,8,12)))
  f[1,1] <- more_id
  
  n <- matrix("",1,13)#96
  
  fx <- rbind(f,n,f1,n,f2) #%>% View()
  colnames(fx) <- 1:dim(fx)[2]
  
  return(fx)
}
```

```{r}
for (i in 1:length(allxls)) {
  path <- allxls[i]
  
  num <- stringr::str_replace(path,"data-raw/Template_ELISA_N(.+).xlsx","\\1")
  fal <- write_plate(path = path,sheet = 1,
                   id_range = "B22:N30",od_range = "B33:N41",
                   more_id = "specie",id = "fal")
  viv <- write_plate(path,1,"B79:N87","B90:N98","specie","viv")
  #fal %>% View()
  #viv %>% View()
  write.table(fal,paste0("data/plater/",num,"fal.csv"),
            row.names = F,col.names = F,sep = ",",quote = F)
  write.table(viv,paste0("data/plater/",num,"viv.csv"),
            row.names = F,col.names = F,sep = ",",quote = F)
}
```

### read_plates

```{r}
a <- "data/plater/11fal.csv"
b <- "data/plater/11viv.csv"
#c <- read.csv(a,header = F) 
#d <- read.csv(b,header = F) 
#c %>% View; d %>% View
plater::check_plater_format(a)
plater::check_plater_format(b)
c <- read_plates(files=c(a,b)) %>% 
  rename_all(funs(stringr::str_to_lower(.))) %>% 
  mutate(plate=stringr::str_replace(plate,"(.+)fal|viv","\\1"))
str(c)
#c %>% dplyr::count(plate)
#c
```

## Output
```{r}
#saveRDS(tbl_df(std.raw), "data/01-standard-raw.rds")
#saveRDS(tbl_df(std.par), "data/01-parametr.rds")
#saveRDS(Pf.ID, "data/01-dilution.rds")
```


***

## APPENDIX: Upper/Lower limits

```{r}
#std.par <- readRDS("data/01-parametr.rds")
std.par
```

***

<!--## References-->
