# ELISA STANDARDIZATION {#stdrad}

## Aim

* Principal: 
    - Implement a reproducible standardization of OD values across ELISA plates following 
    **Miura et.al., 2008**[@Miura2008].

* Secondary: 
    - Generate useful outputs to compare standardization quality.


## To Do

- Implement `plater` and `readxl` data input
- Assess the **inter-plate-CV** comparing STD variability across plates to generate a 
**Limit of Quantification**
    + Actual CV is the **intra-plate-CV**
    + Generate a tibble of all STD dilutions OD's
    + Extend `std.par` output to all parameters
- Other inter-plate measurements:
    + blank, positive and negative controls


## Dependencies

The required R packages for this analysis are:

+ `plater` [@plater]
+ `readxl` [@readxl]
+ `tidyxl` [@tidyxl]
+ `drc` [@Ritz2015]

```{r, results='hide', message=FALSE}
##essential
library(tidyverse)    # set of tidy packages (readr, tibble, dplyr, tidyr, ggplot2, ¿purr?)
library(tidyxl)       # read untidy excel formats
library(readxl)       # read excel files as tidy tables
library(plater)       # read, tidy, display microtiter plates
library(drc)          # fit dose-response models
##accesory
library(DiagrammeR)   # create flowchart
```

```{r}
theme_set(theme_bw())
```


## Method

- using `DiagrammeR` [@DiagrammeR]

```{r, fig.align='center', fig.width=9}
#install.packages("DiagrammeR")
#library(DiagrammeR)

DiagrammeR("
  graph LR
    A[XLS<br>data] -.-> |<center>readxl<br>+tidyxl</center>| B{<center>CSV<br>data<br>w/<br>PLATER<br>format</center>}
    B -.-> |plater| C{<center>TIDY<br>data</center>}
    C --> C1[STD]
    C --> E[ctr +/-]
    C --> D1[UNK]
    
    C1 -.-> |drc| C2[4pLL model]
    C2 --> C3[Box-Cox]
    C3 --> F{<center>UNK<br>Ab.units</center>}

    D1 --> D2[mean.OD]
    D2 --> D3[OD %CV]
    D3 --> F
    
    F --> G1[Histogram]
    F --> G2[Density]
    F --> G3[QQPlot]

    style C fill:#ffffff, stroke:#000000, stroke-width:2px    
    style A fill:#ffffff, stroke:#000000, stroke-width:2px
    style B fill:#ffffff, stroke:#000000, stroke-width:2px
    style C1 fill:#ffffff, stroke:#000000, stroke-width:2px
    style C2 fill:#ffffff, stroke:#000000, stroke-width:2px
    style C3 fill:#ffffff, stroke:#000000, stroke-width:2px
    style D1 fill:#ffffff, stroke:#000000, stroke-width:2px
    style D2 fill:#ffffff, stroke:#000000, stroke-width:2px
    style D3 fill:#ffffff, stroke:#000000, stroke-width:2px
    style E fill:#ffffff, stroke:#000000, stroke-width:2px
    style F fill:#ffffff, stroke:#000000, stroke-width:2px
    style G1 fill:#ffffff, stroke:#000000, stroke-width:2px
    style G2 fill:#ffffff, stroke:#000000, stroke-width:2px
    style G3 fill:#ffffff, stroke:#000000, stroke-width:2px
")
```


### Log-Logistic (4pLL) model

The curve of the **log-logistic symetric** model describe the *response* `f(x)` dependent of the *dose* `x` 
and **04 parameters**: $$ f(x)=f(x;b,c,d,e)=c+\frac{d-c}{1+\exp[b(log(x)-log(e))]}\ $$ where: 
  
- `c` is the **lower limit** of the response when the *dose* `x` approaches infinity, 
- `d` is the **upper limit** when the *dose* `x` approaches zero,
- `b` is the **slope** around the **point of inflection**, represented by 
- `e` defined as **effective dose** and commmonly denoted as [@Ritz2015]:
    + `ED50`, `EC50` or `IC50` for continuous responses,
    + `LD50` or `LC50` for binomial responses, and
    + $T_{50}$ for event-time responses.

## plater

```{r,eval=FALSE,echo=FALSE}
a <- "data/example-1.csv"
b <- read.csv(a,header = F) 
#b %>% View
plater::check_plater_format(a)
c <- read_plate(a)
str(c)
```

```{r,eval=FALSE,echo=FALSE}
a <- "data/example-2-part-A.csv"
d <- "data/example-2-part-B.csv"
b <- read.csv(a,header = T) 
e <- read.csv(d,header = F)
#b %>% View; e %>% View
#plater::check_plater_format(a)
plater::check_plater_format(d)
c <- add_plate(data=b,file=d,well_ids_column = "Wells")
str(c)
view_plate(data = c, well_ids_column = "Wells", 
  columns_to_display = c("Concentration", "Killing"))
```

```{r,eval=FALSE,echo=FALSE}
a <- "data/example-1.csv"
b <- "data/more-bacteria.csv"
c <- read.csv(a,header = F) 
d <- read.csv(b,header = F) 
#c %>% View; d %>% View
plater::check_plater_format(a)
plater::check_plater_format(b)
c <- read_plates(files=c(a,b))
str(c)
c %>% dplyr::count(Plate)
```

## Procedure

### generate list

```{r}
# lista de muestras en el directorio
smp <- list.files("data-raw/") #length(smp)
#smp %>% as.tibble() %>% rename("sample"=value)

# número de muestras por comunidad
tmp <- smp %>% as.tibble() %>% dplyr::rename("sample"=value) %>% 
  mutate(template=stringr::str_replace(sample,"(Template_ELISA).+","ok")) %>% 
  filter(template=="ok")
#tmp$sample

# loop para generar rutas
allxls <-  NULL
for(i in 1:length(tmp$sample)){
  #a <- list.files(paste0("data-raw/",tmp$sample[i]))
  b <- paste0("data-raw/",tmp$sample[i]#,"/",a
              )
  allxls <- c(allxls,b)
}
```

### write_plate

```{r}
write_plate <- function(path,sheet,id_range,od_range,more_id,id) {
  f1 <- readxl::read_xlsx(path,range = id_range,
                        sheet = sheet,col_names = F) %>% as.matrix()
  f2 <- readxl::read_xlsx(path,range = od_range,
                        sheet = sheet,col_names = F) %>% as.matrix()
  
  f1[1,1] <- "id"; f2[1,1] <- "od"; f2[1,-1] <- 1:12
  
  f <- rbind(f1[1,],cbind(f1[-1,1],matrix(id,8,12)))
  f[1,1] <- more_id
  
  n <- matrix("",1,13)#96
  
  fx <- rbind(f,n,f1,n,f2) #%>% View()
  colnames(fx) <- 1:dim(fx)[2]
  
  return(fx)
}
```

```{r}
for (i in 1:length(allxls)) {
  path <- allxls[i]
  
  num <- stringr::str_replace(path,"data-raw/Template_ELISA_N(.+).xlsx","\\1")
  fal <- write_plate(path = path,sheet = 1,
                   id_range = "B22:N30",od_range = "B33:N41",
                   more_id = "specie",id = "fal")
  viv <- write_plate(path,1,"B79:N87","B90:N98","specie","viv")
  #fal %>% View()
  #viv %>% View()
  write.table(fal,paste0("data/plater/",num,"fal.csv"),
            row.names = F,col.names = F,sep = ",",quote = F)
  write.table(viv,paste0("data/plater/",num,"viv.csv"),
            row.names = F,col.names = F,sep = ",",quote = F)
}
```

### read_plates

```{r,eval=FALSE,echo=FALSE}
a <- "data/plater/23fal.csv"
b <- "data/plater/23viv.csv"
#c <- read.csv(a,header = F) 
#d <- read.csv(b,header = F) 
#c %>% View; d %>% View
plater::check_plater_format(a)
plater::check_plater_format(b)
c <- read_plates(files=c(a,b)) %>% 
  rename_all(funs(stringr::str_to_lower(.))) %>% 
  mutate(plate=stringr::str_replace(plate,"(.+)fal|viv","\\1"))
str(c)
#c %>% dplyr::count(plate)
#c
```

```{r}
# lista de muestras en el directorio
ptr <- list.files("data/plater",full.names = T) #length(smp)
```

```{r}
all <- plater::read_plates(files=ptr) %>% 
  rename_all(funs(stringr::str_to_lower(.))) %>% 
  mutate(temp=stringr::str_replace(plate,"(.+)fal|viv","\\1")) %>% 
  separate(wells,c("row","col"),sep = 1,remove = F) %>% 
  mutate(type=ifelse(row=="A" | id=="Blank","std",
                     ifelse(id=="C+" | id=="C-","ctr","unk")),
         ab=ifelse(row=="A",stringr::str_replace(id,"STD 1/(.+)","\\1"),
                    ifelse(id=="Blank","0",NA_character_))
         ) %>% 
  mutate(ab=as.numeric(ab)) %>% 
  mutate(ab=ifelse(row=="A",max(ab,na.rm=T)/ab,ab)) %>% #select(-row,-loc)
  filter(!is.na(id)) %>% 
  #replace_na(list(ID = "na")) %>% 
  #filter(ID!="na") %>% 
  mutate(order=seq(1,dim(.)[1])) #%>% 
  #mutate(specie=as.factor(specie)) %>% 
  #unite(plate,c("plate","specie"),sep = "_",remove = T)
```

```{r,eval=FALSE,echo=FALSE}
all %>% group_by(plate) %>% dplyr::count()
all %>% filter(type=="unk") %>% dplyr::count(plate)
```

### standarization

```{r}
end <- all
# mod is mean od (plus sd and cv)
mod <- end %>% 
  filter(type=="unk") %>% 
  group_by(plate,#anti,
           id) %>% 
  summarise_at(vars(od),c("mean","sd")) %>% 
  ungroup() %>% 
  dplyr::rename(mean.od="mean",
                sd.od="sd") %>% 
  mutate(cv.od=100*sd.od/mean.od) %>% 
  mutate(order=seq(1,dim(.)[1]))
  #filter(ID=="1570")

mab <- end %>% 
  filter(type=="unk") %>% 
  group_by(plate,#anti,
           id) %>% 
  slice(1) %>% 
  ungroup() %>% #filter(ID=="1570")
  mutate(order=seq(1,dim(.)[1])) %>% 
  full_join(mod %>% dplyr::select(order,mean.od,sd.od,cv.od),
            by="order") %>% #mutate(test.id= ID.x==ID.y) %>% dplyr::count(test.id)
  dplyr::select(-order,-od) %>% 
  mutate(ord=seq(1,dim(.)[1])) %>% 
  unite(code,id,ord,sep="_",remove = F)
```

```{r,eval=FALSE, echo=FALSE}
mab %>% dplyr::count(plate)
```

#### blank issue

```{r}
#mascara
blk <- end %>% 
  filter(id=="Blank") %>% 
  group_by(plate) %>% slice(1) %>% ungroup() %>% #
  dplyr::select(-od)
#media por par de blancos por placa
std <- end %>% 
  filter(id=="Blank") %>% 
  group_by(plate) %>% #
  summarise_at(vars(od),mean,na.rm=T) %>% 
  ungroup() %>% 
  #dplyr::rename(mean.OD="OD") %>% 
  full_join(blk,by="plate") %>% 
  dplyr::select(plate,id,type,ab,od,everything()) %>% 
  union(end %>% filter(type=="std" & id!="Blank")) %>% 
  arrange(plate,ab) %>% 
  mutate(plate=as.factor(plate))
```

#### 4pll per template

```{r}
# no phe (from phenoptype)
phe <- std
```

```{r,message=FALSE}

mab_ir <- NULL
mod_bx <- NULL

#
# new dose levels as support for the line
#mdo$abs %>% summary()
new_x <- expand.grid(exp(seq(log(0.1),log(2048),length=100)))
# db to add predictions of all plates
new <- data_frame(ord=as.character(),
                  resp=as.double(),
                  p=as.double(),
                  pmin=as.double(),
                  pmax=as.double(),
                  plate=as.character())
#

for (j in 1:length(levels(phe$plate))) {
  
#
# 5 PARAMETER ESTIMATION 4pLL model
#
wb.m1 <- drm(od ~ ab, plate, 
               data= std %>% filter(plate==levels(phe$plate)[j]),#j
             #data= std,
               fct = LL.4(names = c("b", "c", "d", "e")))
#
wb.model <- wb.m1
# 6 BOX-COX TRANSFORMATION against RESIDUAL heterogeneity
wb.model.BX <- boxcox(wb.model, 
                     main=expression("Optimal " ~ lambda ~ " with confidence intervals"), 
                     plotit = FALSE)
#coefficients(wb.model.BX) %>% matrix(7,4)
mab_p <- mab %>% as.data.frame()
# 7 UNK abS ESTIMATION by INVERSE REGRESSION
mir <- ED(wb.model.BX, 
                 mab_p[mab_p$plate==levels(phe$plate)[j],"mean.od"],#j
                   #wb_MEAN[1:n,5],
                   type = "absolute",interval = "delta",
                   #clevel = "Pfal", 
                 display = FALSE)
  
mab_ir <- rbind(mab_ir,mir)
mod_bx <- rbind(mod_bx,coefficients(wb.model.BX))

#
# predictions and confidence intervals
pdm <- predict(wb.model.BX, newdata = new_x, interval = "confidence")
# new data with predictions
new_p <- bind_cols(new_x %>% 
                     as.tibble() %>% 
                     rownames_to_column(var = "ord")
                   , pdm %>% 
                     as.tibble() %>% 
                     rownames_to_column(var = "ord")
                   ) %>%
  dplyr::select(-ord1) %>% 
  mutate(plate=levels(phe$plate)[j]) %>% 
  dplyr::rename(resp=Var1,p=Prediction,pmin=Lower,pmax=Upper)

new <- union(new,new_p)
#

}

#mab
#mab[mab$plate==levels(phe$plate)[1],] #%>% duplicated() %>% sum()

# 7.1 FEED UNK abS DATA.FRAME
mdo <- mab_ir %>% as.data.frame() %>% rownames_to_column() %>% 
  #dplyr::rename(ord=rowname) %>% 
  separate(rowname,c("par","plate","mean.od.c"),sep = ":") %>% 
  rownames_to_column("ord") %>% 
  #mutate(ord=seq(1,dim(.)[1])) %>% 
  full_join(mab %>% 
              mutate(ord=as.character(ord)) %>% 
              mutate(mean.od.c=as.character(mean.od))
            ,
            by = c("ord","plate","mean.od.c")) %>% 
  dplyr::select(plate,ord,id,code,type,#pheno,igg,
                mean.od,sd.od,cv.od,abs=Estimate,
         everything(),-par,-mean.od.c,-ab) #%>% 
  # ANY MANUAL FILTERING of replicate on different templates could be applied HERE!!!
  #filter(!code=="2235_137") 

mod_bt <- mod_bx %>% as.data.frame() %>% rownames_to_column() %>% as.tibble() %>% 
  full_join(std %>% 
              dplyr::count(plate) %>% dplyr::select(-n) %>% rownames_to_column(),
            by="rowname") %>% 
  dplyr::select(plate,everything(),-rowname) %>% 
  separate(plate,c("temp","specie"),sep = 2,remove = F) #%>% 
  #dplyr::rename(plate=rowname) %>% 
  #mutate(plate=stringr::str_replace(plate,"(\\d)","N\\1"))

new <- new %>% mutate(ord=as.numeric(ord)) %>% arrange(plate,ord)
```

#### outputs

```{r}
#
#fin
#end
#mod
#mab

# standard curve data
std

# estimated ab unit data
mdo

# estimated parameters per standard curve
mod_bt

# predicted model per standard curve 
new
```

#### *CHECKPOINT
```{r}
mdo %>% arrange(id,plate)
#samples with replicates in other PLATE
mdo %>% group_by(id) %>%  dplyr::count() %>% arrange(desc(n)) %>% filter(n!=2) 
mdo %>% filter(id=="LL055-7") %>% arrange(id)
mdo %>% filter(id=="LL028-2") %>% arrange(id)

mdo %>% filter(id=="ZG181-1" | id=="ZG182-1") %>% arrange(id)
mdo <- mdo %>% mutate(id=if_else(id=="ZG182-1","ZG181-1",id))
mdo %>% filter(id=="ZG181-1" | id=="ZG182-1") %>% arrange(id)
```

### quality control plots

```{r}
#std
ctr <- end %>% filter(type=="ctr")
#ctr %>% filter(plate==levels(phe$plate)[1] & id=="C+") %>% .$od
#std %>% filter(plate==levels(phe$plate)[1] & id=="Blank") %>% .$od
```

#### cv

```{r, fig.height=10, fig.width=10}
mdo %>% 
  ggplot(aes(mean.od,cv.od,colour=plate)) +
  geom_point() +
  coord_cartesian(ylim = c(0,100)) +
  geom_hline(aes(yintercept=20),linetype="dashed",size=0.3) +
  geom_vline(aes(xintercept=0.25),linetype="dashed",size=0.3) +
  facet_wrap(~plate,ncol = 7) +
  labs(title="QC plot: Intra-plate coefficient of variation")
```

- __OBSERVACIONES:__
    + La baja variabilidad permite descartar error de pipeteo.
    + Todas las réplicas incluidas en las muestras del estudio tienen un CV menor a 22%.

#### 4pll

```{r, fig.height=10, fig.width=11}
std %>% 
  mutate(ab=ifelse(ab==0,0.5,ab)) %>% 
  ggplot(aes(ab,od)) +
  geom_hline(aes(yintercept=od,linetype=id),data=ctr) +
  geom_point(aes(colour=plate)) +
  #geom_ribbon(data=new, aes(x=resp, y=p, ymin=pmin, ymax=pmax), alpha=0.2) +
  #geom_line(data=new, aes(x=resp, y=p, colour=plate)) +
  scale_x_log10() +#coord_trans(x="log") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~plate,ncol = 7,scales = "free"
             ) +
  labs(title="QC plot: 4-parameter log-logistic model per plate")
  #xlab("Ferulic acid (mM)") + ylab("Root length (cm)")
```

- __OBSERVACIONES:__
    + El cruce del control positivo y la tercera dilución de la curva estandar descartan error en el pipeteo. También indican que cualquier otra anomalía es producto a un error sistemático.
    + Valores de curva estandar por debajo del control negativo anulan la validez de las lecturas por debajo de este valor de od.
    + Entre las posibles fuentes de error están:
        1. Efecto de el uso de un buffer sustrato distinto entre el grupo de placas __N1-N7__ (__OPD__ x60min, todas placas con curva defectuosa) y __N8-N18__ (__TMB__ x60min, todas placas con curva adecuadas).
        2. Alteraciones en la ejecución del protocolo, debido a que las ELISA son muy sensibles a los tiempos de incubación y correcto lavado entre estos.
        3. Ruido de fondo en las placas, por deficiencias en el lavado o uso de placas recicladas.
        4. Diferencias en el accionar entre distintos operadores por placa.

```{r, fig.height=8.1, fig.width=11}
mdo %>% 
  ggplot(aes(abs,mean.od,colour=plate)) +
  #coord_cartesian(ylim = c(0,1)) +
  geom_point() +
  geom_errorbarh(aes(xmin=Lower,xmax=Upper), colour="black", size=.2) +
  scale_x_log10() +
  facet_wrap(~plate#+pheno
             ,ncol = 7,scales = "free"
             ) + 
  labs(title="Estimates of Arbitrary Antibody units (AU) per plate and phenotype") +
  geom_hline(aes(yintercept=od,linetype=id),data=ctr) ## filter threshold?
  # inverse regression method
  #facet_wrap(igg~pheno,nrow = 2)
```

- __OBSERVACIONES:__
    + Lecutas de muestras serán descartadas si tienen:
        - `mean.od` por debajo del cruce con su respectivo control negativo.
        
        <!--- Error estandar mayor a 100.
        + Ver resultado [aqui](#post-quality).-->
### filtering

#### filter replicates

```{r}
rept <- mdo %>% group_by(id,specie) %>% 
  dplyr::count(id) %>% arrange(desc(n)) %>% 
  filter(n>1) %>% dplyr::select(-n) #replicates
#rept

rept_sel <- mdo %>% 
  inner_join(rept,by = c("id","specie")) %>% 
  arrange(id,specie) %>% 
  group_by(id,specie) %>% 
  filter(cv.od == min(cv.od)) # SELECT with the lowest CV
# all in plate5 with lower CV -but- high AU uncertainty ???????
rept_sel

length(unique(mdo$id))
dim(mdo)

mdo <- mdo %>% 
  anti_join(rept,by = c("id")) %>% 
  union(rept_sel) %>% 
  mutate(ord=as.numeric(ord)) %>% 
  arrange(plate,ord)

length(unique(mdo$id))
dim(mdo)
#mdo
#mdo %>% dplyr::count(id,anti) %>% arrange(desc(n))
```

#### *CHECKPOINT

```{r}
mdo %>% filter(id=="LL055-7") %>% arrange(id)
mdo %>% filter(id=="LL028-2") %>% arrange(id)
mdo %>% filter(id=="ZG181-1" | id=="ZG182-1") %>% arrange(id)
```

#### filter NaN

```{r}
### any NaN??

nan <- mdo %>% 
  filter(abs=="NaN") # only 7
nan

#mod_bt[10:13,c(1,2,4,5)]

mdo <- mdo %>% filter(!abs=="NaN")
```

#### *CHECKPOINT

```{r}
mdo
```

### distribution plots

#### linear

##### scale free

```{r,fig.width=6,fig.height=2.2}
##### ab units
mdo %>% 
  ggplot(aes(x=abs#,fill=pheno
             ))+
  geom_histogram(aes(x=abs,..density..),
                 alpha=.5,position = "identity") +
  geom_density(alpha=.5,position = "identity") +
  facet_wrap(~specie,scales = "free")

#ggsave("data-raw/raw/unap-tesis/RAFAEL-data/ab_dens.png")
#Rmisc::multiplot(b,c,cols = 1)
```

#### log

##### scale free

```{r,fig.width=6,fig.height=2.2}
##### ab units
mdo %>% 
  ggplot(aes(x=abs#,fill=pheno
             ))+
  scale_x_log10() +
  geom_histogram(aes(x=abs,..density..),
                 alpha=.5,position = "identity") +
  geom_density(alpha=.5,position = "identity") +
  facet_wrap(~specie,scales = "free")

#ggsave("data-raw/raw/unap-tesis/RAFAEL-data/ab_dens.png")
#Rmisc::multiplot(b,c,cols = 1)
```



## Output
```{r}
#saveRDS(tbl_df(std.raw), "data/01-standard-raw.rds")
#saveRDS(tbl_df(std.par), "data/01-parametr.rds")
#saveRDS(Pf.ID, "data/01-dilution.rds")
```


***

## APPENDIX: Upper/Lower limits

```{r}
#std.par <- readRDS("data/01-parametr.rds")
#std.par
```

***

<!--## References-->
