---
title: "PROJECT: ELISA standardization"
author: "Andree Valle Campos"
date: '`r Sys.Date()`'
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::html_document2:
    keep_md: yes
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    code_folding: "hide"
  #bookdown::pdf_book: default
bibliography: analysis/SeroMarker.bib
biblio-style: apalike
link-citations: yes
csl: analysis/american-medical-association.csl
description: "This is a project notebook using the bookdown package."
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,     # CHANGE TO FALSE IN PDF_DOCUMENTS
                      warning = FALSE) 
#knitr::opts_knit$set(root.dir = '../.')
options(width = 110) # expand limits of CONSOLE output  # CHANGE TO DEFAULT in PDF
# FOR PDF CHANGE: html_document, toc_float, number_sections, echo, width
# FOR PDF add new page USE: \newpage
```

```{r, results='hide', message=FALSE, eval=TRUE, include=FALSE}
require(Hmisc) # to call markupSpecs
mu <- markupSpecs$html # expand limits of HTML SCREEN output to use an entire wide screen
# with the following hidden command (<code>r mu$widescreen()</code>).
#mu$widescreen()
```
`r mu$widescreen()`

***

# Project notebook {-}

This is a project notebook using `Rmarkdown` [@rmarkdown], `knitr` [@knitr], 
`Hmisc` [@Hmisc] and `bookdown` [@bookdown] packages. 
This *dynamic document* format integrates **text**, **code** and **results** 
[@CienciaReproducible2016].

***

## Project Objectives {-}

* __Main:__ 
    - __Standardize__ the OD~450nm~ values across ELISA plates 
    by the estimation of Antibody Units of each unknown sample.
    - __Unificate__ and __Update__ the epidemiological dataset.
    - Generate a __QC report__ of UNAP vs NAMRU measurements. [here](#qc-report)

* __Secondary:__
    - Implement a [reproducible workflow](https://www.ncbi.nlm.nih.gov/pubmed/26776185) 
    for standardization of ELISA plates from a template matrix and subsequent analysis.

***

## Last update {-}

- **[Data Dictionary](#data-dictionary)** available for __epidemiology+serology__ covariates.
- Database of __epidemiology+serology__ covariates in `.dta` format __without labels__. [issue submited](#write-dta)
- `slopegraph` showing the changes of `age` along [updates](#slope). 
- All results at the beginning + __hiperlinks__ of sections __3, 4, 5__.

## Urgent To do's {-}

- CHECK IF **ZG182-1** should be corrected to **ZG181-1**.
    + Missing reads are complementary between ID's.
    + details in [Appendix](#appendix-residual-na) **NOTE 2**
    
- Assess the **inter-plate-CV** comparing STD variability across plates to generate a 
**Limit of Quantification**
    + Actual CV w.r.t OD is the **intra-plate-CV**
    + Generate a tibble of all STD dilutions OD's
    + Extend `std.par` output to all parameters
    + Include other inter-plate measurements:
        + blank, positive and negative controls

***
