---
title: "PROJECT: ELISA standardization"
author: "Andree Valle Campos"
date: '`r Sys.Date()`'
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::html_document2:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
    code_folding: "hide"
  #bookdown::pdf_book: default
bibliography: analysis/SeroMarker.bib
biblio-style: apalike
link-citations: yes
csl: analysis/american-medical-association.csl
description: "This is a project notebook using the bookdown package."
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,     # CHANGE TO FALSE IN PDF_DOCUMENTS
                      warning = FALSE) 
#knitr::opts_knit$set(root.dir = '../.')
options(width = 110) # expand limits of CONSOLE output  # CHANGE TO DEFAULT in PDF
# FOR PDF CHANGE: html_document, toc_float, number_sections, echo, width
# FOR PDF add new page USE: \newpage
```

```{r, results='hide', message=FALSE, eval=TRUE}
require(Hmisc) # to call markupSpecs
mu <- markupSpecs$html # expand limits of HTML SCREEN output to use an entire wide screen
# with the following hidden command (<code>r mu$widescreen()</code>).
#mu$widescreen()
```
`r mu$widescreen()`

# Project notebook {-}

This is a project notebook using the bookdown package.

## Last update

- **[Data Dictionary](#data-dictionary)** available
- Data epidemiological covariates output as a `.dta` without data labels. [reasons](#write-dta)

## Urgent To do's

- CHECK IF **ZG182-1** should be corrected to **ZG181-1**.
    + Missing reads are complementary between ID's.
    
- Assess the **inter-plate-CV** comparing STD variability across plates to generate a 
**Limit of Quantification**
    + Actual CV w.r.t OD is the **intra-plate-CV**
    + Generate a tibble of all STD dilutions OD's
    + Extend `std.par` output to all parameters
    + Include other inter-plate measurements:
        + blank, positive and negative controls

## README file

```{bash}
less README.md
```

